---
title: "Time Series Forecast"
author: "Uche Kalu"
date: "4/15/2022"
output:
  html_document: default
  pdf_document: default
---

### The Goal of this Analysis is to predict the future as accurately as possible based on current information available .

**When forecasting time series data, the aim is to estimate how the sequence of observation will continue into the future**

```{r message=FALSE, warning=FALSE}
library(modeltime)   
library(tidymodels)   
library(tidyverse)    
library(timetk)
library(lubridate)   
library(glmnet)
```

### DATA

The Dataset we will be using is [**bike_sharing_daily**]{.ul}. It is a built-in dataset from timetk package.

We will be using only two variables from the dataset - dteday and cnt columns. cnt variable is the response variable

```{r results='hide'}
View(bike_sharing_daily)    

mydata <- bike_sharing_daily %>%
  select(dteday, cnt)    

mydata
```

**This displays an interactive time series chart using the stated variables**

```{r fig.width=11, fig.height=6}
mydata %>% plot_time_series(dteday, cnt)
```

### TRAIN / TEST SPLITS

We will split the dataset into training and testing data by using the last 3 months in the data as testing set

```{r warning=FALSE, message=FALSE}
splits <- time_series_split(mydata, assess = "3 months",
                            cumulative = TRUE) 
```

**Visualizing the training and testing data**

```{r fig.width=12, fig.height=6}
splits %>%
  tk_time_series_cv_plan() %>%
  plot_time_series_cv_plan(dteday, cnt)  

```

### Creating FORECAST using [Auto-Arima]{.ul} model, [Prophet]{.ul} model and [GLM]{.ul} machine learning algorithm

#### AUTO ARIMA model (this is an auto-regressive forecasting algorithm)

```{r message=FALSE, results='hide'}
mymodel <- arima_reg() %>%
  set_engine("auto_arima") %>%
  fit(cnt ~ dteday, training(splits))   

```

#### Facebook Prophet model

```{r message=FALSE, results='hide'}
ourmodel <- prophet_reg(seasonality_yearly = TRUE) %>%
  set_engine("prophet") %>%
  fit(cnt ~ dteday, training(splits))

```

#### GLM Machine Learning algorithm

```{r message=FALSE, results='hide'}
yrmodel <- linear_reg(penalty = 0.01) %>%
  set_engine("glmnet") %>%
  fit(cnt ~ wday(dteday, label = TRUE) +
        month(dteday, label = TRUE) +
        as.numeric(dteday), training(splits))
```

### Using MODELTIME package to organize and compare our models

**Create a Modeltime Table (this helps to organize our models)**

```{r}
hismodel <- modeltime_table(mymodel, ourmodel, yrmodel)  
hismodel
```

### Calibration.

**Here we calculate predictions and residuals(error) for the test data**

```{r}
mycalib <- hismodel %>%
  modeltime_calibrate(testing(splits))  
mycalib
```

### Viewing Accuracy.

**Lets see the accuracy from our testing dataset predictions**

ae - Mean Absolute Error (is the average error aggregated for across d prediction. d smaller d value, d better); we see that ARIMA model is d worst

rsq - R-Squared value (Its value is from 0 to 1. the higher d value, d better): we see PROPHET model is d best

```{r}
mycalib %>% modeltime_accuracy() 
```

### Visualization of the Testing data

**Displays an interactive time series chart using the three models on training and testing data**

We see that ARIMA model is not following d time series trend unlike the other models so you can disable ARIMA in d legend bar to remove it from the chart

```{r fig.width=12, fig.height=6}
mycalib %>%
  modeltime_forecast(new_data = testing(splits),
                     actual_data = mydata) %>%  
  plot_modeltime_forecast()
```

### Forecast The Future

**We will forecast for the next 3 months**

We re-train the three models on the full dataset (not the training data) so we get the most accurate predictions in the future

```{r message=FALSE, results='hide'}
myforecast <- mycalib %>%
  modeltime_refit(mydata) %>%      
  modeltime_forecast(h = "3 months", actual_data = mydata)   

```

```{r fig.height=6, fig.width=12}
myforecast %>%
  plot_modeltime_forecast()
```
